
# Examples

Here is an example of the tables produced by {epibuffet} from a randomly
generated data set based on an MSF data dictionary for Measles data. 

```{r example}
library(epidict)
library(matchmaker)
library(epikit)
library(dplyr)
library(epibuffet)

linelist <- epidict::gen_data("Measles", numcases = 1000, org = "MSF")
measles_dict <- epidict::msf_dict("Measles", compact = FALSE)

# Cleaning linelist data
linelist_clean <- matchmaker::match_df(
  x = linelist,
  dictionary = measles_dict,
  from = "option_code",
  to = "option_name",
  by = "data_element_shortname",
  order = "option_order_in_set"
)
linelist_clean
```

## Quick tabulations

```{r tab_linelist}
the_symptoms <- tab_linelist(linelist_clean,
  cough, nasal_discharge, severe_oral_lesions,
  transpose = "value"
) 
the_symptoms
```

```{r print_linelist}
the_symptoms %>%
  epikit::rename_redundant("%" = "proportion") %>%
  epikit::augment_redundant("(n)" = "n") %>%
  knitr::kable()
```

# 2x2 tables

In R, creating 2x2 tables is as simple as using the function `table()`, but
unfortunately, it can be difficult to interpret the values of these tables
because the dimensions are often flipped around for different analyses. The
function `data_frame_from_2x2()` will present these values labeled unambiguously.

```{r symp}
symptoms_tf <- linelist_clean %>%
  dplyr::transmute(
    pneumonia = grepl("Yes", pneumonia),
    cough = grepl("Yes", cough),
    nasal_discharge = grepl("Yes", nasal_discharge),
    oral_lesions = grepl("Yes", severe_oral_lesions),
    contact = grepl("known case", contact_history)
  ) 

symptoms_tf

print(pxc <- with(symptoms_tf, table(pneumonia, cough)))
print(pxcxc <- with(symptoms_tf, table(pneumonia, cough, contact)))

data_frame_from_2x2(pxc)
data_frame_from_2x2(pxcxc)
```


## Odds / Risk / Incidence Rate Ratios

```{r orrr}
tu <- tab_univariate(symptoms_tf, 
  outcome = pneumonia, 
  cough, nasal_discharge, oral_lesions, contact,
  mergeCI = TRUE,
  extend_output = FALSE,
  measure = "OR"
)
tu %>% 
  select(-est_type) %>% 
  epikit::augment_redundant("exposed " = "exp_") %>%
  rename(odds = est_ci) %>%
  knitr::kable(digits = 3)
```


